%
% Esperanta gramatiko por angloj
%
\label{gram:angla}
\markright{GRAMMAR}
\thispagestyle{plain}
\begin{center}
\phantomsection
\narrow{\LARGE\bf \spaceoutless{GRAMMAR}}
\addcontentsline{toc}{section}{Grammar (Gramatiko Angla)}
\selectlanguage{english}

\rule{13mm}{0.4pt}
\vspace{2em}

{\large\gramsec{A) THE ALPHABET}}
\vspace{1em}

\setstretch{1}
\begin{tabu} to \textwidth{+Y@{}ZY@{}ZY@{}ZY@{}ZY@{}ZY@{}ZY}
\rowstyle{\Large\arbfont} Aa, & Bb, & Cc, & Ĉĉ, & Dd, & Ee, & Ff, \\
\rowstyle{\footnotesize} \emph{a} as in „last“ & \emph{b} as in „be“ & \emph{ts} as in „wits“ & \emph{ch} as in „church“ & \emph{d} as in „do“ & \emph{e} as in „make“ & \emph{f} as in „fly“ \\[1ex]
\rowstyle{\Large\arbfont} Gg, & Ĝĝ, & Hh, & Ĥĥ, & Ii, & Jj, & Ĵĵ, \\
\rowstyle{\footnotesize} \emph{g} as in „gun“ & \emph{j} as in „join“ & \emph{h} as in „half“ & strongly aspirated h, \emph{ch} as in „loch“ (Scotch) & \emph{i} as in „marine“ & \emph{y} as in „yoke“ & \emph{z} as in „azure“  \\[1ex]
\rowstyle{\Large\arbfont} Kk, & Ll, & Mm, & Nn, & Oo, & Pp, & Rr, \\
\rowstyle{\footnotesize} \emph{k} as in „key“ & \emph{l} as in „line“ & \emph{m} as in „make“ & \emph{n} as in „now“ & \emph{o} as in „not“ & \emph{p} as in „pair“ & \emph{r} as in „rare“ \\[1ex]
\rowstyle{\Large\arbfont} Ss, & Ŝŝ, & Tt, & Uu, & Ŭŭ, & Vv, & Zz. \\
\rowstyle{\footnotesize}  \emph{s} as in „see“ & \emph{sh} as in „show“ & \emph{t} as in „tea“ & \emph{u} as in „bull“ & \emph{u} as in „mount“ (used in diphthongs) & \emph{v} as in „very“ & \emph{z} as in „zeal“ 
\end{tabu}
\end{center}

{\footnotesize {\bf Remark.} — If it be found impractical to print works with the diacritical signs (ˆ,˘), the letter \emph{h} may be substituted for the sign (ˆ), and the sign ( ˘ ) may be altogether omitted.}
\begin{center}
\large \gramsec{B) PARTS OF SPEECH}
\end{center}

\textbf{1.} There is no indefinite, and only one definite, article, \emph{la}, for all genders, numbers, and cases.

\textbf{2.} Substantives are formed by adding \emph{o} to the root. For the plural, the letter \emph{j} must be added to the singular. There are two cases: the nominative and the objective (accusative). The root with the added \emph{o} is the nominative, the objective adds an \emph{n} after the \emph{o}. Other cases are formed by prepositions; thus, the possessive (genitive) by \emph{de}, “of”; the dative by \emph{al}, “to”; the instrumental (ablative) by \emph{kun}, “with”, or other preposition as the sense demands. E.~g., root \emph{patr}, “father”; \emph{la patr$'$o}, “the father”; \emph{patr$'$o$'$n}, “father” (objective), \emph{de la patr$'$o}, “of the father”, \emph{al la patr$'$o}, “to the father”, \emph{kun la patr$'$o}, “with the father”; \emph{la patro$'$j}, “the fathers”; \emph{la patro$'$j$'$n}, “the fathers” (obj.), \emph{por la patr$'$o$'$j}, “for the fathers”.

\textbf{3.} Adjectives are formed by adding \emph{a} to the root. The numbers and cases are the same as in substantives. The comparative degree is formed by prefixing \emph{pli} (more); the superlative by \emph{plej} (most). The word “than” is rendered by \emph{ol}, e.~g., \emph{pli blank$'$a ol neĝ$'$o}, “whiter than snow”.

\textbf{4.} The cardinal numerals do not change their forms for the different cases. They are :
\emph{unu} (1), \emph{du} (2), \emph{tri} (3), \emph{kvar} (4), \emph{kvin} (5), \emph{ses} (6), \emph{sep} (7), \emph{ok} (8), \emph{naŭ} (9), \emph{dek} (10), \emph{cent} (100), \emph{mil} (1000).

The tens and hundreds are formed by simple junction of the numerals, e.~g., 583\eraro{533} = \emph{kvin$'$cent tri$'$dek tri}.

Ordinals are formed by adding the adjectival \emph{a} to the cardinals, e.~g., \emph{unu$'$a}, “first”; \emph{du$'$a}, “second”, etc.

Multiplicatives (as “threefold”, “fourfold”, etc.) add \emph{obl}, e.~g., \emph{tri$'$obl$'$a}, “threefold”.

Fractionals add \emph{on}, as \emph{du$'$on$'$o}, “a half”, \emph{kvar$'$on,o}, “a quarter”. Collective numerals add \emph{op}, as \emph{kvar$'$op$'$e}, “four together”.

Distributives prefix \emph{po}, e.~g., \emph{po kvin}, “five apiece”.

Adverbials take \emph{e}, e.~g., \emph{unu$'$e}, “firstly”, etc.

\textbf{5.} The Personal Pronouns are: \emph{mi}, I; \emph{vi}, thou, you; \emph{li}, he; \emph{ŝi}, she; \emph{ĝi}, it; \emph{si}, “self”; \emph{ni}, “we”; \emph{ili}, “they”; \emph{oni}, “one”, “people”, (French “on”).

Possessive pronouns are formed by suffixing to the required personal, the adjectival termination. The declension of the pronouns is identical with that of substantives. E.~g., \emph{mi}, “I”; \emph{mi$'$n}, “me” (obj.); \emph{mi$'$a}, “my”, “mine”.

\textbf{6.} The verb does not change its form for numbers or persons, e.~g., \emph{mi far$'$as}, “I do”; \emph{la patr$'$o far$'$as}, “the father does”; \emph{ili far$'$as}, “they do”.

\begin{center}
\bf Forms of the Verb:
\end{center}

a) The present tense ends in \emph{as}, e.~g., \emph{mi far$'$as}, “I do”.

b) The past tense ends in \emph{is}, e.~g., \emph{li far$'$is}, “he did”.

c) The future tense ends in \emph{os}, e.~g., \emph{ili far$'$os}, “they will do”.

ĉ) The subjunctive mood ends in \emph{us}, e.~g., \emph{ŝi far$'$us}, “the\eraro{``she''} may do”.

d) The imperative mood ends in \emph{u}, e.~g., \emph{ni far$'$u}, “let us do”.

e) The infinitive mood ends in \emph{i}, e.~g., \emph{far$'$i}, “to do”.

There are two forms of the participle in the international language, the changeable or adjectival, and the unchangeable or adverbial. 

f) The present participle active ends in \emph{ant}, e.~g., \emph{far$'$ant$'$a}, “he who is doing”; \emph{far$'$ant$'$e}, “doing”.

g) The past participle active ends in \emph{int}, e.~g., \emph{far$'$int$'$a}, “he who has done”; \emph{far$'$int$'$e}, “having done”.

ĝ) The future participle active ends in \emph{ont}, e.~g., \emph{far$'$ont$'$a}, “he who will do”; \emph{far$'$ont$'$e}, “about to do”.

h) The present participle passive ends in \emph{at}, e.~g., \emph{far$'$at$'$e}, “being done”.

ĥ) The past participle passive ends in \emph{it}, e.~g., \emph{far$'$it$'$a}, “that which has been done”; \emph{far$'$it$'$e}, “having been done”.

i) The future participle passive ends in \emph{ot}, e.~g., \emph{far$'$ot$'$a}, “that which will be done”; \emph{far$'$ot$'$e}, “about to be done”.

All forms of the passive are rendered by the respective forms of the verb \emph{est} (to be) and the present participle passive of the required verb; the preposition used is \emph{de}, “by”. E.~g., \emph{ŝi est$'$as am$'$at$'$a de ĉiu$'$j}, “she is loved by every one.”

\textbf{7.} Adverbs are formed by adding \emph{e} to the root. The degrees of comparison are the same as in adjectives, e.~g., \emph{mi$'$a frat$'$o kant$'$as pli bon$'$e ol mi}, “my brother sings better than I”.

\textbf{8.} All prepositions govern the nominative case.

\begin{center}
\large \gramsec{C) GENERAL RULES}
\end{center}

\textbf{9.} Every word is to be read exactly as written, there are no silent letters.

\textbf{10.} The accent falls on the last syllable but one, (penultimate).

\textbf{11.} Compound words are formed by the simple junction of roots, (the principal word standing last), which are written as a single word, but, in elementary works, separated by a small line ($'$). Grammatical terminations are considered as independent words, e.~g., \emph{vapor$'$ŝip$'$o}, “steamboat”, is composed of the roots \emph{vapor}, “steam”, and \emph{ŝip}, “a boat”, with the substantival termination \emph{o}.

\textbf{12.} If there be one negative in a clause, a second is not admissible.

\textbf{13.} In phrases answering the question “where?” (meaning direction), the words take the termination of the objective case; e.~g., \emph{kie$'$n vi ir$'$as?} “where are you going?” \emph{dom$'$o$'$n}, “home”; \emph{London$'$o$'$n}, “to London”; etc.

\textbf{14.} Every preposition in the international language has a definite fixed meaning. If it be necessary to employ some preposition, and it is not quite evident from the sense which it should be, the word \emph{je} is used, which has no definite meaning; for example, \emph{ĝoj$'$i je tio}, “to rejoice \emph{over} it”; \emph{rid$'$i je tio} “to laugh \emph{at} it”; \emph{enu$'$o je la patr$'$uj$'$o}, “a longing \emph{for} one’s fatherland”. In every language different prepositions, sanctioned by usage, are employed in these dubious cases, in the international language, one word, \emph{je}, suffices for all. Instead of \emph{je}, the objective without a preposition may be used, when no confusion is to be feared.

\textbf{15.} The so-called “foreign” words, i.~e., words which the greater number of languages have derived from the same source, undergo no change in the international language, beyond conforming to its system of orthography.---Such is the rule with regard to primary words, derivatives are better formed (from the primary word) according to the rules of the international grammar: e.~g., \emph{teatr$'$o}, “theater”, but \emph{teatr$'$a}, “theatrical”, (not \emph{teatrical$'$a}), etc.

\textbf{16.} The \emph{a} of the article, and the final \emph{o} of substantives, may be sometimes dropped euphoniae gratia, e.~g., \emph{de l’ mond$'$o} for \emph{de la mond$'$o}; \emph{Ŝiller’} for \emph{Ŝiller$'$o}; in such cases an apostrophe should be substituted for the discarded vowel. 

\newpage