% Internacia Lingvo (de D-ro Esperanto)
% alinome Unua Libro
%
% de D-ro L. L. ZAMENHOF
% angla traduko de Richard H. GEOGHEGAN, 1889
%
% XeLaTeX-a versio de Shawn C. KNIGHT, 2019
%
% This work is licensed under a Creative Commons 
% Attribution-NonCommercial-ShareAlike 4.0 International License.
% The original work by Dr. Zamenhof is in the public domain.
%
% Ĉi tiu verko estas permesita per Creative Commons 
% Attribution-NonCommercial-ShareAlike 4.0 Internacia Permesilo.
% La originala verko de D-ro Zamenhof estas senkopirajta.
%
\documentclass[12pt,twoside]{book}

\input{enkonduko.tex}
%
% End preamble and then begin title page
% 
\begin{document}
\sloppy
\begin{titlepage}

\newlength{\tempparskip}
\setlength{\tempparskip}{\parskip}
\setlength{\parskip}{2ex}
\newlength{\tempparindent}
\setlength{\tempparindent}{\parindent}

\vspace*{\fill}

\begin{center}
\scalebox{0.8}[1]{\latinia{\LARGE D\doktoror{} ESPERANTO’S}}

\scalebox{1}[3]{\didone{\LARGE INTERNATIONAL LANGUAGE,}}

{\sansfont{\LARGE INTRODUCTION}}

{\Large \&}

\scalebox{1.8}[1.8]{\shadowtext{\cowboyfont{\LARGE Complete\hspace{1ex}Grammar.}}}

\poranglojbox

\scalebox{0.45}[1]{\copper{\huge ENGLISH EDITION}}

{\small BY}

\scalebox{1.5}[0.7]{\cowboyfont{\LARGE R. H. Geoghegan}}

\scalebox{1}[0.7]{\monastic{\large Balliol College, Oxford.}}\\[1ex]
\pgfornament[width=0.4\textwidth]{89}

\scalebox{0.8}[1]{\large \angloj{WARSAW.}}

\scalebox{1.5}[0.7]{\cowboyfont{\Large L. Samenhof, Przejazd N. 9.}}

\rule{3em}{0.4pt}

\scalebox{1.5}[1]{{1889.}}

\vspace*{\fill}

\end{center}

\setlength{\parskip}{\tempparskip}
\end{titlepage}

%
% Inside title page
%
\renewcommand{\footrulewidth}{0.4pt}
\setstretch{1.1}

\vspace*{12em}

\begin{table}[h]
\begin{center}
{ДОЗВОЛЕНО ЦЕНЗУРОЮ \\
\vspace{1ex}
Варшава, 5 Января 1889 года.}
\end{center}

\end{table}

\vspace{5em}

{\small \finger An international language, like every national one, is the property of society, and the author renounces all personal rights in it forever.}

\vspace{5em}

\begin{flushright}
\begin{tabular}{r}
\hline
\bf\footnotesize Printed by Ch. Kelter Nowolipie Str. N. 11.
\end{tabular}
\end{flushright}

\setlength{\parskip}{0pt}

\vspace*{\fill}

%
% Introduction
%
\titleformat{\chapter}[display]{\centering}{\chaptertitlename}{0pt}{\Huge}
\renewcommand{\footrulewidth}{0pt}
\chapter*{\scalebox{0.8}[1]{INTRODUCTION.}}
\fancyhead[C]{--- \thepage{} ---}

\begin{center}
\rule{5em}{0.4pt}
\end{center}

The reader will doubtless take up this little work with an incredulous smile, supposing that he is about to peruse the impracticable schemes of some good citizen of Utopia. I would, therefore, in the first place, beg of him to lay aside all prejudice, and treat seriously and critically the question brought before him.

I need not here point out the considerable importance to humanity of an international language,---a language unconditionally accepted by everyone, and the common property of the whole world. How much time and labour we spend in learning foreign tongues, and yet when travelling in foreign countries, we are, as a rule, unable to converse with other human beings in their own language. How much time, labour, and money are wasted in translating the literary productions of one nation into the language of another, and yet, if we rely on translations alone, we can become acquainted with but a tithe of foreign literature.

Were there but an international language, all translations would be made into it alone, as into a tongue intelligible to all, and works of an international character would be written in it in the first instance.

The Chinese wall dividing literatures would disappear, and the works of other nations would be as readily intelligible to us as those of our own authors. Books being the same for everyone, education, ideals, convictions, aims, would be the same too, and all nations would be united in a common brotherhood. Being compelled, as we now are, to devote our time to the study of several different languages, we cannot study any of them sufficiently well, and there are but few persons who can even boast a complete mastery of their mother-tongue; on the other hand, languages cannot progress towards perfection, and we are often obliged, even in speaking our own language, to borrow words and expressions from foreigners, or to express our thoughts inexactly.

How different would the case be, had we but two languages to learn; we should know them infinitely better, and the languages themselves would grow richer, and reach a higher degrees of perfection than is found in any of those now existing. And yet, though language is the prime motor of civilisation, and to it alone we owe the having raised ourselves above the level of other animals, difference of speech is a cause of antipathy, nay even of hatred, between people, as being the first thing to strike us on meeting. Not being understood we keep aloof, and the first notion that occurs to our minds is, not to find out whether the others are of our own political opinions, or whence their ancestors came from thousands of years ago, but to dislike the strange sound of their language. Any one, who has lived for a length of time in a commercial city, whose inhabitants were of different unfriendly nations, will easily understand what a boon would be conferred on mankind by the adoption of an international idiom, which, without interfering with domestic affairs or the private-life of nations, would play the part of an official and commercial dialect, at any rate in countries inhabited by people of different nationalities.

The immense importance, which it may well be imagined, an international language would acquire in science, commerce, etc., I will not here expatiate on: whoever has but once bestowed a thought on the subject will surely acknowledge that no sacrifice would be too great, if by it we could obtain a universal tongue. It is, therefore, imperative that the slightest effort in that direction should be attended to. The best years of my life have been devoted to the momentous cause which I am now bringing before the public, and I hope that, on account of the importance of the subject, my readers will peruse this pamphlet attentively to the end.

I shall not here enter upon an analysis of the various attempts already made to give the public a universal language, but will content myself with remarking that these efforts have amounted, either to a short system of mutually-intelligible signs, or to a natural simplification of the grammar of existing modern languages, with a change of their words into arbitrarily-formed ones. The attempts of the first category were quickly seen to be too complicated for practical use, and so faded into oblivion; those of the second were, perhaps, entitled to the name of “languages”, but certainly not “international” languages. The inventors called their tongues “universal”, I know not why, possibly, because no one in the whole world except themselves could understand a single word, written or spoken in any of them. If a language, in order to become universal, has but to be named so, then, forsooth, the wish of any single individual can frame out of any existing dialect a universal tongue. As these authors na\"{i}vely imagined that their essays would be enthusiastically welcomed and taken up by the whole world, and as this unanimous welcome is precisely what the cold and indifferent world declines to give, when there is no chance of realising any immediate benefit, it is not much to be marvelled at, if these brilliant attempts came to nothing. The greater part of the world was not in the slightest degree interested in the prospect of a new language, and the persons who really cared about the matter thought it scarcely worth while to learn a tongue which none but the inventor could understand. When the whole world, said they, has learnt this language, or at least several million people, we will do the same. And so a scheme, which had it but been able to number some thousands of adepts before its appearance in public, would have been enthusiastically hailed, came into the world an utter fiasco. If the \glqq{}Volapük\grqq{}, one of the latest attempts at a universal tongue, has indeed its adepts, it owes its popularity solely to the idea of its being a “universal language”, and that idea has in itself something so attractive and sublime, that true enthusiasts, leaders in every new discovery, are ready to devote their time, in the hope that they may, perchance, win the cause.

But the number of enthusiasts, after having risen to a certain number, will remain stationary\footnote{One cannot, of course, reckon the number of those who learned the language as equal to the number of instruction-books sold.} and as the unfeeling and indifferent world will never consent to take any pains in order to speak with the few, this attempt will, like its predecessors, disappear without having achieved any practical victory.

I have always been interested in the question of a universal language, but as I did not feel myself better qualified for the work than the authors of so many other fruitless attempts, I did not risk running into print, and merely occupied myself with imaginary schemes and a minute study of the problem. At length, however, some happy ideas, the fruits of my reflections, incited me to further work, and induced me to essay the systematic conquest of the many obstacles, which beset the path of the inventor of a new rational universal language. As it appears to me that I have almost succeeded in my undertaking, I am now venturing to lay before the critical public, the results of my long and assiduous labours.

The principal difficulties to be overcome were:

1) To render the study of the language so easy as to make its acquisition mere play to the learner.

2) To enable the learner to make direct use of his knowledge with persons of any nationality, whether the language be universally accepted or not; in other words, the language is to be directly a means of international communication.

3) To find some means of overcoming the natural indifference of mankind, and disposing them, in the quickest manner possible, and \emph{en masse}, to learn and use the proposed language as a living one, and not only in last extremities, and with the key at hand.

Amongst the numberless projects submitted at various times to the public, often under the high-sounding but unaccountable name of “universal languages”, no one has solved at once more than \inbold{one} of the above-mentioned problems, and even that but partially. (Many other problems, of course, presented themselves, in addition to those here noticed, but these, as being of but secondary importance, I shall not in this place discuss).

Before proceeding to enlighten the reader as to the means employed for the solution of the problems, I would ask of him to reconsider the exact significance of each separately, so that he may not be inclined to carp at my methods of solution, merely because they may appear to him perhaps too simple. I do this, because I am well aware that the majority of mankind feel disposed to bestow their consideration on any subject the more carefully, in proportion as it is enigmatical and incomprehensible. Such persons, at the sight of so short a grammar, with rules so simple, and so readily intelligible, will be ready to regard it with a contemptuous glance, never considering the fact, --- of which a little further reflection would convince them, --- that this simplification and bringing of each detail out of its original complicated form into the simplest and easiest conceivable, was, in fact, the most insuperable obstacle to be coped with. 

\vspace{12pt}

{\hfil \scalebox{2}[1]{\large\cowboyfont I.}}

\vspace{12pt}

The first of the problems was solved in the following manner:

1) I simplified the grammar to the utmost, and while, on the one hand, I carried out my object in the spirit of the existing modern languages, in order to make the study as free from difficulties as possible, on the other hand I did not deprive it of clearness, exactness, and flexibility. My whole grammar can be learned perfectly in \inbold{one hour}. The immense alleviation given to the study of a language, by such a grammar, must be self-evident to everyone.

2) I established rules for the formation of new words, and at the same time, reduced to a very small compass the list of words absolutely necessary to be learned, without, however, depriving the language of the means of becoming a rich one. On the contrary, thanks to the possibility of forming from one root-word any number of compounds, expressive of every conceivable shade of idea, I made it the richest of the rich amongst modern tongues. This I accomplished by the introduction of numerous prefixes and suffixes, by whose aid the student is enabled to create new words for himself, without the necessity of having previously to learn them, e.~g.

1) The prefix \emph{mal} denotes the direct opposite of any idea. If, for instance, we know the word for “good”, \emph{bon\,a}, we can immediately form that for “bad”, \emph{mal\,bon\,a}, and hence the necessity of a special word for “bad” is obviated. In like manner, \emph{alt\,a}, “high”, “tall”, \emph{mal\,alt\,a,} “low”, “short”; \emph{estim\,i,} “to respect”, \emph{mal\,estim\,i}, “to despise”, etc. Consequently, if one has learned this single word \emph{mal} he is relieved of leaning a long string of words such as “hard” (premising that he knows “soft”), “cold”, “old”, “dirty”, “distant”, “darkness”, “shame”, “to hate”, etc., etc.

2) The suffix \emph{in} marks the feminine gender, and thus if we know the word “brother”, \emph{frat\,o}, we can form “sister”, \emph{frat\,in\,o}: so also, “father”, \emph{patr\,o}; “mother”, \emph{patr\,in\,o}. By this device words like “grandmother”, “bride”, “girl”, “hen”, “cow”, etc., are done away with.

3) The suffix \emph{il} indicates an instrument for a given purpose, e.~g., \emph{tranĉ\,i}, “to cut”, \emph{tranĉ\,il\,o}, “a knife”; so words like “comb”, “axe”, “bell”, etc., are rendered unnecessary.

In the same manner are employed many other affixes, --- some fifty in all, --- which the reader will find in the vocabulary at end of this tractate.\footnote{To facilitate the finding of these affixes they are entered in the vocabulary as separate words.} Moreover, as I have laid it down as a general rule, that every word already regarded as international, --- the so-called “foreign” words, for example, --- undergoes no change in my language, except such as may be necessary to bring it into conformity with the international orthography, innumerable words become superfluous, e.~g., “locomotive”, “telegraph”, “nerve”, “temperature”, “centre”, “form”, “public”, “platinum”, “figure”, “waggon”, “comedy”, and hundreds more.

By the help of these rules, and others, which will be found in the grammar, the language is rendered so exceedingly simple that the whole labour in learning consists in committing to memory some 900 words, --- which number includes all the grammatical inflexions, prefixes, etc. --- With the assistance of the rules given in the grammar, any one of ordinary intellectual capacity, may form for himself all the words, expressions, and idioms in ordinary use. Even these 900 words, as will be shown directly, are so chosen, that the learning them offers no difficulty to a well-educated person.

Thus the acquirement of this rich, mellifluous, universally-comprehensible language, is not a matter of years of laborious study, but the mere light amusement of a few days. 

\vspace{12pt}

{\hfil \scalebox{2}[1]{\large\cowboyfont II.}}

\vspace{12pt}

The solution of the second problem was effected thus:

1) I introduced a complete dismemberment of ideas into independent words, so that the whole language consists, not of words in different states of grammatical inflexion, but of unchangeable words. If the reader will turn to one of the pages of this book written in my language, he will perceive that each word always retains its original unalterable form, ---namely, that under which it appears in the vocabulary. The various grammatical inflexions, the reciprocal relations of the members of a sentence, are expressed by the junction of immutable syllables. But the structure of such a synthetic language being altogether strange to the chief European nations, and consequently difficult for them to become accustomed to, I have adapted this principle of dismemberment to the spirit of the European languages, in such a manner that anyone learning my tongue from grammar alone, without having previously read this introduction, --- which is quite unnecessary for the learner, --- will never perceive that the structure of the language differs in any respect from that of his mother-tongue. So, for example, the derivation of \emph{frat\,in\,o}, which is in reality a compound of \emph{frat} “child of the same parents as one’s self”, \emph{in} “female”, \emph{o} “an entity”, “that which exists”, i.~e., “that which exists as a female child of the same parents as one’s self” = “a sister”,--- is explained by the grammar thus: the root for “brother” is \emph{frat}, the termination of substantives in the nominative case is \emph{o}, hence \emph{frat\,o} is the equivalent of “brother”; the feminine gender is formed by the suffix \emph{in}, hence \emph{frat\,in\,o} = “sister”. (The little strokes, between certain letters, are added in accordance with a rule of the grammar, which requires their insertion between each component part of every complete word). Thus the learner experiences no difficulty, and never even imagines that what he calls terminations, suffixes, etc.,---are complete and independent words, which always keep their own proper significations, whether placed at the beginning or end of a word, in the middle, or alone. The result of this construction of the language is, that everything written in it can be immediately and perfectly understood by the help of the vocabulary, --- or even almost without it, --- by anyone who has not only not learnt the language before, but even has never heard of its very existence. Let me illustrate this by an example: --- I am amongst Englishmen, and have not the slightest knowledge of the English language; I am absolutely in need of making myself understood, and write in the international tongue, may be, as follows:

    \emph{Mi ne sci\,as ki\,e mi las\,is la baston\,o\,n; ĉu vi ĝi\,n ne vid\,is?}

I hold out to one of the strangers an International--English vocabulary, and point to the title, where the following sentence appears in large letters: “Everything written in the international language can be translated by the help of this vocabulary. If several words together express but a single idea, they are written as one word, but separated by commas; e.~g., \emph{frat\,in\,o}, though a single idea is yet composed of three words which must be looked for separately in the vocabulary”. If my companion has never heard of the international language he will probably favour me at first with a vacant stare, will then take the paper offered to him, and searching for the words in the vocabulary, as directed, will make out something of this kind:

\smaller

\begingroup
\renewcommand{\arraystretch}{1}
\begin{longtable}{>{\itshape}c@{ }l@{ }>{\itshape}l>{\ttfamily}l@{ }p{4cm}@{ }l@{ }c}
Mi & \tlba & mi & = & I & \trba & I \\
ne & \tlba & ne & = & not & \trba & not \\
\multirow{2}{*}{sci\,as} & \tlbb & sci & = & know & \trbb & \multirow{2}{*}{do know} \\
 & & as & = & sign of the present tense \\
kie & \tlba & kie & = & where & \trba & where \\
mi & \tlba & mi & = & I & \trba & I \\
\multirow{2}{*}{las\,is} & \tlbb & las & = & leave & \trbb & \multirow{2}{*}{have left} \\
 & & is & = & sign of the past tense \\
la & \tlba & la & = & the & \trba & the \\
\multirow{3}{*}{baston\,o\,n;} & \tlbc & baston & = & stick & \trbc & \multirow{3}{*}{stick;} \\
 & & o & = & sign of a substantive \\
 & & n & = & sign of the objective case \\
\multirow{2}{*}{ĉu} & \tlbb & ĉu & = & {whether, if, employed in questions} & \trbb & \multirow{2}{*}{whether} \\
vi & \tlba & vi & = & you, thou & \trba & you \\
\multirow{2}{*}{ĝi\,n} & \tlbb & ĝi & = & it, this & \trbb & \multirow{2}{*}{it} \\
 & & n & = & sign of the objective case \\
ne & \tlba & ne & = & not & \trba & not \\
\multirow{2}{*}{vid\,is?} & \tlbb & vid & = & see & \trbb & \multirow{2}{*}{have seen?} \\
 & & is & = & sign of the past tense \\
\end{longtable}
\endgroup
\normalsize
And thus the Englishman will easily understand what it is I desire. If he wishes to reply, I show him an English--International vocabulary, on which are printed these words: “To express anything by means of this vocabulary, in the international language, look for the words required, in the vocabulary itself; and for the terminations necessary to distinguish the grammatical forms, look in the grammatical appendix, under the respective headings of the parts of speech which you desire to express”. Since the explanation of the whole grammatical structure of the language is comprised in a few lines,---as a glance at the grammar will show,---the finding of the required terminations occupies no longer time than the turning up a word in the dictionary.

I would now direct the attention of my readers to another matter, at first sight a trifling one, but, in truth, of immense importance. Everyone knows the impossibility of communicating intelligibly with a foreigner, by the aid of even the best of dictionaries, if one has no previous acquaintance with the language. In order to find any given word in a dictionary, we must know its derivation, for when words are arranged in sentences, nearly every one of them undergoes some grammatical change. After this alteration, a word often bears not the least resemblance to its primary form, so that without knowing something of the language beforehand, we are able to find hardly any of the words occurring in a given phrase, and even those we do find will give no connected sense. Suppose, for example, I had written the simple sentence adduced above, in German: \glqq Ich weiss nicht wo ich den Stock gelassen habe; haben Sie ihn nicht gesehen?\grqq{} Anyone who did not speak or understand German, after searching for each word separately in a dictionary, would produce the following farrago of nonsense: “I; white; not; where; I; --- ; stick; dispassionate; property; to have; she, they, you; --- ; not; --- ?” I need scarcely point out that a lexicon of a modern language is usually a tome of a certain bulk, and the search for any number of words one by one is in itself a most laborious undertaking, not to speak of the different significations attaching to the same word amongst which there is but a bare possibility of the student selecting the right one. The international vocabulary, owing to the highly synthetic structure of the language, is a mere leaflet, which one might carry in one’s note-book, or the waistcoat-pocket. Granted that we \emph{had} a language with a grammar simplified to the utmost, and whose every word had a definite fixed meaning, the person addressed would require not only to have beforehand some knowledge of the grammar, to be able, even with the vocabulary at hand, to understand anything addressed to him, but would also need some previous acquaintance with the vocabulary itself, in order to be able to distinguish between the primitive word and its grammatically-altered derivatives. The utility, again, of such a language would wholly depend upon the number of its adepts, for when sitting, for instance, in a railway-carriage, and wishing to ask a fellow-traveller, “How long do we stop at ---?”, it is scarcely to be expected that he will undertake to learn the grammar of the language before replying! By using, on the other hand, the international language, we are set in possibility of communicating directly with a person of any nationality, even though he may never have heard of the existence of the language before.

Anything whatever, written in the international tongue, can be translated, without difficulty, by means of the vocabulary alone, no previous study being requisite. The reader may easily convince himself of the truth of this assertion, by experimenting for himself with the specimens of the language appended to this pamphlet. A person of good education will seldom need to refer to the vocabulary, a linguist scarcely at all.

Let us suppose that you have to write to a Spaniard, who neither knows your language nor you his. You think that probably he has never heard of the international tongue. --- No matter, write boldly to him in that language, and be sure he will understand you perfectly. The complete vocabulary required for everyday use, being but a single sheet of paper, can be bought for a few pence, in any language you please, easily enclosed in the smallest envelope, and forwarded with your letter. The person to whom it is addressed will without doubt understand what you have written, the vocabulary being not only a clue to, but a complete explanation of your letter. The wonderful power of combination possessed by the words of the international language renders this lilliputian lexicon amply sufficient for the expression of every want of daily life; but words seldom met with, technical terms, and foreign words familiar to all nations, as, “tobacco”, “theatre”, “fabric”, etc., are not included in it. If such words, therefore, are needed, and it is impossible to express them by some equivalent terms, the larger vocabulary must be consulted.

2) It has now been shown how, by means of the peculiar structure of the international tongue, any one may enter into an intelligible correspondence with another person of a different nationality. The sole drawback, until the language becomes more widely known, is the necessity under which the writer is placed of waiting until the person addressed shall have analysed his thoughts. In order to remove this obstacle, as far as practicable, at least for persons of education, recourse was had to the following expedient. Such words as are common to the languages of all civilised peoples, together with the so-called “foreign” words, and technical terms, were left unaltered. If a word has a different sound in different languages, that sound has been chosen which is common to at least two or three of the most important European tongues, or which, if found in one language only, has become familiar to other nations. When the required word has a different sound in every language, some word was sought for, having only a relative likeness in meaning to the other, or one which, though seldom used, is yet well-known to the leading nations, e.~g., the word for ``near'' is different in every European language, but if one consider for a moment the word ``proximus'' (nearest), it will be noticed that some modified form of the word is in use in all important tongues. If, then, I call ``near'', \emph{proksim}, the meaning will be apparent to every educated man. In other emergencies words were drawn from the Latin, as being a quasi-international language. Deviations from these rules were only made in exceptional cases, as for the avoidance of homonyms, simplicity of orthography, etc. In this manner, being in communication with a European of fair education, who has never learnt the international tongue, one may make sure of being immediately understood, without the person addressed having to refer continually to the vocabulary.

In order that the reader may prove for himself the truth of all that has been set forth above, a few specimens of the international language are subjoined.\footnote{In correspondence with persons who have learnt the language, as well as in works written for them exclusively, the commas, separating parts of words, are omitted.}


\secthead{Patr\,o ni\,a.}{2ex}{2ex}

Patr\,o ni\,a, kiu est\,as en la ĉiel\,o, sankt\,a est\,u Vi\,a nom\,o, ven\,u reĝ\,ec\,o Vi\,a, est\,u vol\,o Vi\,a, kiel en la ĉiel\,o, tiel ankaŭ sur la ter\,o. Pan\,o\,n ni\,a\,n ĉiu\,tag\,a\,n don\,u al ni hodiaŭ, kaj pardon\,u al ni ŝuld\,o\,j\,n ni\,a\,j\,n, kiel ni ankaŭ pardon\,as al ni\,a\,j ŝuld\,ant\,o\,j; ne konduk\,u ni\,n en tent\,o\,n; sed liber\,ig\,u ni\,n de la mal\,ver\,a, ĉar Vi\,a est\,as la reg\,ad\,o, la fort\,o, kaj la glor\,o etern\,e. Amen!


\secthead{El la Bibli\,o.}{2ex}{2ex}

Je la komenc\,o Di\,o kre\,is la ter\,o\,n kaj la ĉiel\,o\,n. Kaj la ter\,o est\,is sen\,form\,a kaj dezert\,a, kaj mal\,lum\,o est\,is super la profund\,aĵ\,o, kaj la anim\,o de Di\,o si\,n port\,is super la akv\,o. Kaj Di\,o dir\,is: est\,u lum\,o; kaj far\,iĝ\,is lumo. Kaj Di\,o vid\,is la lum\,o\,n ke ĝi est\,as bon\,a, kaj nom\,is Di\,o la lum\,o\,n tag\,o, kaj la mal\,lum\,o\,n Li nom\,is nokt\,o. Kaj est\,is vesper\,o, kaj est\,is maten\,o --- unu tag\,o. Kaj Di\,o dir\,is: est\,u firm\,aĵ\,o inter la akv\,o, kaj ĝi apart\,ig\,u akv\,o\,n de akv\,o. Kaj Di\,o kre\,is la firm\,aĵ\,o\,n kaj apart\,ig\,is la akv\,o\,n kiu est\,as sub la firm\,aĵ\,o, de la akv\,o kiu est\,as super la firm\,aĵ\,o; kaj far\,iĝ\,is tiel. Kaj Di\,o nom\,is la firm\,aĵ\,o\,n ĉiel\,o. Kaj est\,is vesper\,o, kaj est\,is maten\,o --- la du\,a tag\,o. Kaj Di\,o dir\,is: kolekt\,u si\,n la akv\,o de sub la ĉiel\,o unu lok\,o\,n, kaj montr\,u si\,n sek\,aĵ\,o; kaj far\,iĝ\,is tiel. Kaj Di\,o nom\,is la sek\,aĵ\,o\,n ter\,o, kaj la kolekt\,oj\,n de la akv\,o Li nom\,is mar\,o\,j.

\secthead{Leter\,o.}{2ex}{2ex}

\hspace{5em} Kar\,a amik\,o!

Mi prezent\,as al mi kia\,n vizaĝ\,o\,n vi far\,os post la ricev\,o de mi\,a leter\,o. Vi rigard\,os la sub\,skrib\,o\,n kaj ek\,kri\,os: “ĉu li perd\,is la saĝ\,o\,n? Je kia lingv\,o li skrib\,is? Kio\,n signif\,as la foli\,et\,o, kiu\,n li al\,don\,is al si\,a leter\,o?” Trankvil\,iĝ\,u, mi\,a kar\,a! Mi\,a saĝ\,o, kiel mi almenaŭ kred\,as, est\,as tut\,e en ord\,o.

Mi leg\,is antaŭ kelk\,a\,j tag\,o\,j libr\,et\,o\,n sub la nom\,o \glqq{}Lingv\,o inter\,naci\,a\grqq{}. La aŭtor\,o kred\,ig\,as, ke per tiu lingv\,o oni pov\,as est\,i kompren\,at\,a de la tut\,a mond\,o, se eĉ la adres\,it\,o ne sol\,e ne sci\,as la lingv\,o\,n, sed eĉ ankaŭ ne aŭd\,is pri ĝi; oni dev\,as sol\,e al\,don\,i al la leter\,o mal\,grand\,a\,n foli\,et\,o\,n nom\,at\,a\,n \glqq{}vort\,ar\,o\grqq{}. Dezir\,ant\,e vid\,i, ĉu tio est\,as ver\,a, mi skrib\,as al vi en tiu lingv\,o, kaj mi eĉ unu vort\,o\,n ne al\,met\,as en ali\,a lingv\,o, tiel kiel se ni tut\,e ne kompren\,us unu la lingv\,o\,n de la ali\,a. Respond\,u al mi, ĉu vi efektiv\,e kompren\,is kio\,n mi skrib\,is. Se la afer\,o propon\,it\,a de la aŭtor\,o est\,as efektiv\,e bon\,a, oni dev\,as per ĉiu\,j fort\,o\,j li\,n help\,i. Kiam mi hav\,os vi\,a\,n respond\,o\,n, mi send\,os al vi la libr\,et\,o\,n; montr\,u ĝi\,n al ĉiu\,j loĝ\,ant\,o\,j de vi\,a urb\,et\,o, send\,u ĝin ĉiu\,n vilaĝ\,o\,n ĉirkaŭ la urb\,et\,o, ĉiu\,n urb\,o\,n kaj urb\,et\,o\,n, kie vi nur hav\,as amik\,o\,j\,n aŭ kon\,at\,o\,j\,n. Est\,as neces\,e ke grand\,eg\,a nombr\,o da person\,o\,j don\,u si\,a\,n voĉ\,o\,n --- tiam post la plej mal\,long\,a temp\,o est\,os decid\,it\,a afer\,o, kiu pov\,as port\,i grand\,eg\,a\,n util\,o\,n al la hom\,a societ\,o.

\secthead{Mi\,a pens\,o.}{2ex}{0pt}

\begin{verse}
        \hspace{1em} Sur la kamp\,o, for de l’mond\,o,\\
        Antaŭ nokt\,o de somer\,o\\
        Amik\,in\,o en la rond\,o\\
        Kant\,as kant\,o\,n pri l’esper\,o.\\
        Kaj pri viv\,o detru\,it\,a\\
        Ŝi rakont\,as kompat\,ant\,e, ---\\
        Mi\,a vund\,o re\,frap\,it\,a\\
        Mi\,n dolor\,as re\,sang\,ant\,e.\\

\pstars

        \hspace{1em} \glqq{}Ĉu vi dorm\,as? Ho, sinjor\,o,\\
        Kial tia sen\,mov\,ec\,o?\\
        Ha, kred\,ebl\,e re\,memor\,o\\
        El la kar\,a infan\,ec\,o?\grqq{}\\
        Kio\,n dir\,i? Ne plor\,ant\,a\\
        Pov\,is est\,i parol\,ad\,o\\
        Kun fraŭl\,in\,o ripoz\,ant\,a\\
        Post somer\,a promen\,ad\,o!\\

\pstars

        \hspace{1em} Mi\,a pens\,o kaj turment\,o,\\
        Kaj dolor\,o\,j kaj esper\,o\,j!\\
        Kiom de mi en silent\,o\\
        Al vi ir\,is jam ofer\,o\,j!\\
        Kio\,n hav\,is mi plej kar\,a\,n ---\\
        \emph{La jun\,ec\,o\,n} --- mi plor\,ant\,a\\
        Met\,is mem sur la altar\,o\,n\\
        De la dev\,o ordon\,ant\,a!\\

\pstars

        \hspace{1em} Fajr\,o\,n sent\,as mi intern\,e,\\
        Viv\,i ankaŭ mi dezir\,as, ---\\
        Io pel\,as mi\,n etern\,e,\\
        Se mi al gaj\,ul\,o\,j ir\,as \ldots{} \\
        Se ne plaĉ\,as al la sort\,o\\
        Mi\,a pen\,o kaj labor\,o ---\\
        Ven\,u tuj al mi la mort\,o,\\
        En esper\,o --- sen dolor\,o!
\end{verse}

\secthead{El Heine’.}{0pt}{0pt}

\begin{verse}
        \hspace{1em} En sonĝ\,o princ\,in\,o\,n mi vid\,is\\
        Kun vang\,o\,j mal\,sek\,a\,j de plor\,o, ---\\
        Sub arb\,o, sub verd\,a ni sid\,is\\
        Ten\,ant\,e si\,n kor\,o ĉe kor\,o.\\

\pstars

        \hspace{1em} \glqq{}De l’patr\,o de l’vi\,a la kron\,o\\
        Por mi ĝi ne est\,as hav\,ind\,a;\\
        For, for li\,a sceptr\,o kaj tron\,o ---\\
        Vi\,n mem mi dezir\,as, am\,ind\,a!\grqq{}\\

\pstars

        \hspace{1em} --- \glqq{}Ne ebl\,e!\grqq{} ŝi al mi re\,dir\,as:\\
        \glqq{}En tomb\,o mi est\,as ten\,at\,a,\\
        Mi nur en la nokt\,o el\,ir\,as\\
        Al vi, mi\,a sol\,e am\,at\,a!\grqq{}\\
\end{verse}

\secthead{Ho, mi\,a kor’.}{0pt}{0pt}

\begin{verse}
\hspace{1em} Ho, mi\,a kor’, ne bat\,u mal\,trankvil\,e. \\
El mi\,a brust\,o nun ne salt\,u for! \\
Jam ten\,i mi\,n ne pov\,as mi facil\,e \\
Ho, mi\,a kor’! \\

\pstars

\hspace{1em} Ho, mi\,a kor’! Post long\,a labor\,ad\,o \\
Ĉu mi ne venk\,os en decid\,a hor’! \\
Sufiĉ\,e! trankvil\,iĝ\,u de l’bat\,ad\,o, \\
Ho, mi\,a kor’! 
\end{verse}
\begin{center}
\sectionline
\vspace{12pt}
\end{center}

\hspace*{\fill}\scalebox{2}[1]{\large\cowboyfont{}III.}\hspace*{\fill}

\vspace{12pt}
\nopagebreak
I have now completed my analysis of the more remarkable features of my international language. I have shown the advantages to be derived from a study of it, and proved that its ultimate success is altogether independent of the opinions that may be formed as to its right to the title “international”. For even should the language never come into general use, it gives to every one who \emph{has} learned it, the possibility of being understood by foreigners, if only they be able to read and write. But my tongue has yet another object; not content with internationality, it aims at universality, and aspires to being \emph{spoken} by the majority of educated people. To count on the aid of the public in a scheme of this nature would indeed be to build on a tottering, --- nay rather, an imaginary, --- foundation. The larger part of the public does not care to aid anyone, it prefers to have its wishes gratified without inconvenience to itself. On this account I made my best endeavours to discover some means of accomplishing my object, independently of the help of the public. One of my plans, of which I shall now speak more at large, is a kind of “universal vote”.

If the reader consider all that has been said above, he must come to the conclusion that the study of the international language is practically useful, and completely remunerates the learner for the small amount of trouble he has to expend on it. For my own part, I am naturally wishful that the whole of mankind should take up my language, but I had rather be prepared for the worst, than form too sanguine anticipations. I suppose therefore, that, just at first, very few will consider my language worth the learning, so far as practical usefulness is concerned, and for abstract principles no one will lose even a single hour.

Most of my readers will, either pay not the slightest attention to my proposition, or, doubting whether the language be of any use, never “screw up their courage to the sticking-point” of learning it, fearing that they may be dubbed “dreamers”, a sobriquet dreaded by most people more than fire. What, then, is to be done, to dispose this mass of indifferent and undecided beings to master the international language? Could we, in imagination, look for a moment into the mind of each of these indifferent ones, we should find their thoughts to be taking somewhat of the following form. In principle, no one has anything to oppose to the introduction of an international dialect; on the contrary, all would give it their fullest approval, but each wishes to see the greater part of the civilized world able to speak the language, and himself able to comprehend it, without any preliminary “wearisome bitterness of learning”, on his own part. \emph{Then}, of course, even the most indifferent would set to work, because to shirk the small amount of labour necessary for learning a language possessed of such valuable qualities, and above all, considered \emph{``the thing''} by all the educated, would be regarded as simple stupidity.

In order to supply a language ready for immediate use, without any one having to initiate the study, and to see on every hand people either already proficient in the tongue, or having promised to take it up, we must proceed somewhat in the following manner. Doubtless this little book will be scattered through various countries, and fall into the hands of various readers. I do not ask any of my readers to spend time, labour, or money on the subject now brought to their notice. I merely beg of you, the present reader of the pamphlet, to take up your pen for a moment, fill in one of the appended \emph{\glqq Promes\,o\,j\grqq} (below) and send it to me (Dr. Esperanto, $\mathrm{^c\!/\!_o}$ Dr. L. Samenhof. Warsaw, Poland). The \emph{\glqq Promes\,o\grqq} is to this effect:

\begin{adjustwidth}{1cm}{1cm}
\hspace{1.5em} “I, the undersigned, promise to learn the international language, proposed by Dr. Es\-per\-anto, if it shall be shown that ten million similar promises have been publicly given”.
\end{adjustwidth}

If you have any objections to make to the present form of the language, strike out the words of the promise, and write \emph{\glqq kontraŭ\grqq} (against), beneath them. If you undertake to learn the language unconditionally, i.~e., without reference to the number of other students, strike out the latter words of the \glqq Promes\,o\grqq, and write \emph{\glqq{}sen\,kondiĉ\,e\grqq{}}, (unconditionally). On the back of the promise write name and address. The signing of this promise lays no obligations upon the person signing, and does not bind him to the smallest sacrifice or work. It merely puts him under an obligation to study the language, when ten million other persons shall be doing the same. When that time arrives, there will be no talking about ``sacrifice'', everyone will be ready to study the language, without having signed any promises.

On the other hand, every person signing one of these \glqq{}Promes\,o\,j\grqq{}, will, --- without any greater inconvenience to himself than dipping a pen in ink, --- be hastening on the realization of the traditional ideal of mankind, the universal language. When the number of promises has reached ten millions, a list of the names of those who have signed will be published, and with it, the question of an international language --- decided.

Nothing actually \emph{prevents} people from inducing their friends and acquaintances to sign a promise in any cause, yet how few, as a fact, ever do sign anything, be the object ever so important and advantageous to mankind. More especially, when, as in the present instance, the act of signing, while contributing to the realization of a sublime ideal, at the same time requires no moral nor material sacrifice, can one see no very clear grounds for a refusal.

Doubtless, no one has anything to say, in general, against the introduction of an international language; but, if anyone does not approve of the present form of the language, by all means let him send me, instead of his “Promise”, his “Protest”. For it is, manifestly, the duty of every person able to read and write, of every age, sex, or profession, to give his opinion in this great undertaking; the more so, as it requires no greater sacrifice than that of a few moments for filling in the promise, and a few pence for sending it to me.

I would here beg of all editors of newspapers and magazines to make known the cause to their readers, and at the same time, I would request \emph{my} readers to mention the subject to all their friends.

I need not say any more. I am not so conceited as to suppose that my language is so perfect as to be incapable of improvement, but I make bold to think that I have satisfied all the conditions required in a language claiming to be styled “international”. It is only after having solved successfully all the problems I had proposed to myself,---concerning the more important of which only, I have been able to speak above, owing to the small compass of this pamphlet,---and after many years spent in a careful study of the subject that I venture to appear in public. I am but human; I may have erred, I may have committed unpardonable faults. I may even have omitted to give to my language the very thing most important to it. For these reasons, before printing complete vocabularies and bringing out books and magazines, I lay my work before the public, for the space of one year, addressing myself to the whole intelligent world with the earnest request to send me opinions on the proposed international language. I invite everyone to communicate with me as to the changes, corrections, etc., which he deems advisable. All such observations sent to me, I will gratefully make use of, if they appear really advantageous, and at the same time, not subversive of the fundamental principles of the structure of the language:---that is to say, simplicity, and adaptability to international communication whether adopted \emph{universally} or not.

At the end of the alloted time, an abstract of the proposed changes will be published and the language will receive its final form. But if, even then, anyone should find the language not altogether satisfactory to himself, he should not forget that the language is by no means proof against all further changes, only that the right of alteration will be no longer the author’s personal privilege, but that of an academy of the tongue.

It is no easy task to invent an international language, but it is a still less easy one to persuade the public to make use of it. Hence, it is of the utmost importance that every possible effort be made for its furtherance. When the form of the language has been decided, and the language itself has come into general use, a special academy can introduce, --- gradually and imperceptibly, --- all necessary changes, even should the result be a total alteration of the form of the language. On this account, I would pray those of my readers, who may be, for whatever reasons, dissatisfied with my language, to send in their protests only in the event of their having serious cause for it, such as the finding in the language objectionable features, unalterable in the future.

This little work, which has cost much labour and health, I now commend to the kindly attention of the public, hoping that all, to whom the public weal is dear, will aid me to the best of their ability. Circumstances will show each one in what way he can be of use; I will only direct the attention of all friends of the international language, to that most important object, towards which all eyes must be turned, the success of the voting. Let each do what he can, and in a short time we shall have, that which men have been dreaming of so long, --- \emph{``A Universal Tongue''}.

\sectionline

\newpage

\emph{NB.} The author requests his reader to fill in one of the ``Promises'' on the following page, and send it to him, and to distribute the others amongst friends and acquaintances for the same purpose.

Author's Address:

\vspace{1ex}

\hspace{1.5em} Dr. Esperanto,

\vspace{1ex}

\hspace{3em} $\mathrm{^c\!/\!_o}$ Dr. L. Samenhof,

\vspace{1ex}

\hspace{4.5em} Warsaw,

\vspace{1ex}

\hspace{6em} Russ-Poland.

\sectionline

\cleardoublepage

\begin{table}[h]
\centering
{\setlength{\extrarowheight}{2ex}%
\begin{tabular}{ccc}
\promeso{} & & \promeso \\
\promeso & & \promeso \\
\end{tabular}}
\end{table}

\begin{table}[h]
\centering
{\setlength{\extrarowheight}{2ex}%
\begin{tabular}{ccc}
\nomadreso & & \nomadreso \\
\nomadreso & & \nomadreso \\
\end{tabular}}
\end{table}


\begin{table}[h]
\centering
{\setlength{\extrarowheight}{2ex}%
\begin{tabular}{ccc}
\promeso & & \promeso \\
\promeso & & \promeso \\
\end{tabular}}
\end{table}

\begin{table}[h]
\centering
{\setlength{\extrarowheight}{2ex}%
\begin{tabular}{ccc}
\nomadreso & & \nomadreso \\
\nomadreso & & \nomadreso \\
\end{tabular}}
\end{table}

%
% Grammar
%
\titlespacing*{\chapter}{0pt}{0pt}{0pt}
\titleformat{\section}[display]{\centering}{\sectiontitlename}{0pt}{\Large}
\chapter*{\scalebox{1.5}[1.5]{\LARGE\shadowtext{\bosox{COMPLETE GRAMMAR}}}}
\begin{center}
\scalebox{0.7}[1]{\LARGE OF THE INTERNATIONAL LANGUAGE.}
\sectionline

\section*{\scalebox{0.8}[1]A. \hspace{.2em} \scalebox{0.85}{\grammarpartsfont{The \hspace{.2em} Alphabet.}}}
\setstretch{1}
\begin{tabularx}{\textwidth}{+Y@{}^Y@{}^Y@{}^Y@{}^Y@{}^Y}
\rowstyle{\LARGE} A a, & B b, & C c, & Ĉ ĉ, & D d, & E e,  \\ 
\rowstyle{\footnotesize} \emph{a} as in ``last'' & \emph{b} as in ``be'' & \emph{ts} as in ``wits'' & \emph{ch} as in ``church'' & \emph{d} as in ``do'' & \emph{e} as in ``make'' \\[3ex] 
\rowstyle{\LARGE} F f, & G g, & Ĝ ĝ, & H h, & Ĥ ĥ, & I i, \\
\rowstyle{\footnotesize} \emph{f} as in ``fly'' & \emph{g} as in ``gun'' & \emph{j} as in ``join'' & \emph{h} as in ``half'' & strongly aspirated h, \emph{ch} as in ``loch'' (Scotch) & \emph{i} as in ``marine''  \\[9ex]
\rowstyle{\LARGE} J j, & Ĵ ĵ, & K k, & L l, & M m, & N n,  \\
\rowstyle{\footnotesize} \emph{y} as in ``yoke'' & \emph{z} as in ``azure'' & \emph{k} as in ``key'' & \emph{l} as in ``line'' & \emph{m} as in ``make'' & \emph{n} as in ``now'' \\[3ex]
\rowstyle{\LARGE}O o, & P p, & R r, & S s, & Ŝ ŝ, & T t,   \\
\rowstyle{\footnotesize} \emph{o} as in ``not'' & \emph{p} as in ``pair'' & \emph{r} as in ``rare'' & \emph{s} as in ``see'' & \emph{sh} as in ``show'' & \emph{t} as in ``tea'' \\[3ex]
\rowstyle{\LARGE} & U u, & Ŭ ŭ, & V v, & Z z. \\
\rowstyle{\footnotesize} & \emph{u} as in ``bull'' & \emph{u} as in ``mount'' (used in diphthongs) & \emph{v} as in ``very'' & \emph{z} as in ``zeal'' 
\end{tabularx}
\end{center}
\setstretch{1.1}

If it be found impractical to print works with the diacritical signs ( ˆ, ˘ ), the letter \emph{h} may be substituted for the sign ( ˆ ), and the sign ( ˘ ) may be altogether omitted; but at the beginning of works so printed there should be this note: ``NB ch \texttt{=} ĉ; gh \texttt{=} ĝ; hh \texttt{=} ĥ; jh \texttt{=} ĵ; sh \texttt{=} ŝ.''

When it is necessary to make use of the ``internal'' sign ( \, ) care should be taken that it can not be mistaken for a comma.  Instead of ( \, ) may be printed ( {\relsize{-2.5}$'$} ) or ( $\cdot$ ), e.~g. \emph{sign\,et\,o, sign{\relsize{-2.5}$'$}et{\relsize{-2.5}$'$}o,} or \emph{sign$\cdot$et$\cdot$o}.

\section*{\scalebox{0.8}[1]B. \hspace{.2em} \scalebox{0.85}{\grammarpartsfont{Parts \hspace{.2em} of \hspace{.2em} Speech.}}}

1. There is no indefinite, and only one definite, article, \emph{la}, for all genders, numbers, and cases.

2. Substantives are formed by adding \emph{o} to the root. For the plural, the letter \emph{j} must be added to the singular. There are two cases: the nominative and the objective (accusative). The root with the added \emph{o} is the nominative, the objective adds an \emph{n} after the \emph{o}. Other cases are formed by prepositions; thus, the possessive (genitive) by \emph{de}, “of”; the dative by \emph{al}, “to”; the instrumental (ablative) by \emph{kun}, “with”, or other preposition as the sense demands. E.~g., root \emph{patr}, “father”; \emph{la patr\,o}, “the father”; \emph{patr\,o\,n}, “father” (objective), \emph{de la patr\,o}, “of the father”, \emph{al la patr\,o}, “to the father”, \emph{kun la patr\,o}, “with the father”; \emph{la patro\,j}, “the fathers”; \emph{la patro\,j\,n}, “the fathers” (obj.), \emph{por la patr\,o\,j}, “for the fathers”.

3. Adjectives are formed by adding \emph{a} to the root. The numbers and cases are the same as in substantives. The comparative degree is formed by prefixing \emph{pli} (more); the superlative by \emph{plej} (most). The word “than” is rendered by \emph{ol}, e.~g., \emph{pli blank\,a ol neĝ\,o}, “whiter than snow”.

4. The cardinal numerals do not change their forms for the different cases. They are:

\begin{center}
\begin{tabular}{rlccrl}
1 & \emph{unu} & \hspace{2em} & \hspace{2em} & 7 & \emph{sep} \\
2 & \emph{du} & & & 8 & \emph{ok} \\
3 & \emph{tri} & & & 9 & \emph{naŭ} \\
4 & \emph{kvar} & & & 10 & \emph{dek} \\
5 & \emph{kvin} & & & 100 & \emph{cent} \\
6 & \emph{ses} & & & 1000 & \emph{mil} 
\end{tabular}
\end{center}

The tens and hundreds are formed by simple junction of the numerals, e.~g., 533 = \emph{kvin\,cent tri\,dek tri}.

Ordinals are formed by adding the adjectival \emph{a} to the cardinals, e.~g., \emph{unu\,a}, “first”; \emph{du\,a}, “second”, etc.

Multiplicatives (as “threefold”, “fourfold”, etc.) add \emph{obl}, e.~g., \emph{tri\,obl\,a}, “threefold”.

Fractionals add \emph{on}, as \emph{du\,on\,o}, “a half”, \emph{kvar\,on,o}, “a quarter”. Collective numerals add \emph{op}, as \emph{kvar\,op\,e}, “four together”.

Distributives prefix \emph{po}, e.~g., \emph{po kvin}, “five apiece”.

Adverbials take \emph{e}, e.~g., \emph{unu\,e}, “firstly”, etc.

5. The Personal Pronouns are: \emph{mi}, I; \emph{vi}, thou, you; \emph{li}, he; \emph{ŝi}, she; \emph{ĝi}, it; \emph{si}, “self”; \emph{ni}, “we”; \emph{ili}, “they”; \emph{oni}, “one”, “people”, (French “on”).

Possessive pronouns are formed by suffixing to the required personal, the adjectival termination. The declension of the pronouns is identical with that of substantives. E.~g., \emph{mi}, “I”; \emph{mi\,n}, “me” (obj.); \emph{mi\,a}, “my”, “mine”.

6. The verb does not change its form for numbers or persons, e.~g., \emph{mi far\,as}, “I do”; \emph{la patr\,o far\,as}, “the father does”; \emph{ili far\,as}, “they do”.

Forms of the Verb:

a) The present tense ends in \emph{as}, e.~g., \emph{mi far\,as}, “I do”.

b) The past tense ends in \emph{is}, e.~g., \emph{li far\,is}, “he did”.

c) The future tense ends in \emph{os}, e.~g., \emph{ili far\,os}, “they will do”.

ĉ) The subjunctive mood ends in \emph{us}, e.~g., \emph{ŝi far\,us}, “she may do”.

d) The imperative mood ends in \emph{u}, e.~g., \emph{ni far\,u}, “let us do”.

e) The infinitive mood ends in \emph{i}, e.~g., \emph{far\,i}, “to do”.

There are two forms of the participle in the international language, the changeable or adjectival, and the unchangeable or adverbial. 

f) The present participle active ends in \emph{ant}, e.~g., \emph{far\,ant\,a}, “he who is doing”; \emph{far\,ant\,e}, “doing”.

g) The past participle active ends in \emph{int}, e.~g., \emph{far\,int\,a}, “he who has done”; \emph{far\,int\,e}, “having done”.

ĝ) The future participle active ends in \emph{ont}, e.~g., \emph{far\,ont\,a}, “he who will do”; \emph{far\,ont\,e}, “about to do”.

h) The present participle passive ends in \emph{at}, e.~g., \emph{far\,at\,e}, “being done”.

ĥ) The past participle passive ends in \emph{it}, e.~g., \emph{far\,it\,a}, “that which has been done”; \emph{far\,it\,e}, “having been done”.

i) The future participle passive ends in \emph{ot}, e.~g., \emph{far\,ot\,a}, “that which will be done”; \emph{far\,ot\,e}, “about to be done”.

All forms of the passive are rendered by the respective forms of the verb \emph{est} (to be) and the present participle passive of the required verb; the preposition used is \emph{de}, “by”. E.~g., \emph{ŝi est\,as am\,at\,a de ĉiu\,j}, “she is loved by every one.”

7. Adverbs are formed by adding \emph{e} to the root. The degrees of comparison are the same as in adjectives, e.~g., \emph{mi\,a frat\,o kant\,as pli bon\,e ol mi}, “my brother sings better than I”.

8. All prepositions govern the nominative case.

\section*{\scalebox{0.8}[1]C. \hspace{.2em} \scalebox{0.85}{\grammarpartsfont{General \hspace{.2em} Rules.}}}

1. Every word is to be read exactly as written, there are no silent letters.

2. The accent falls on the last syllable but one, (penultimate).

3. Compound words are formed by the simple junction of roots, (the principal word standing last), which are written as a single word, but, in elementary works, separated by a small line (\, or {\relsize{-2.5}$'$}). Grammatical terminations are considered as independent words, e.~g., \emph{vapor\,ŝip\,o}, “steamboat”, is composed of the roots \emph{vapor}, “steam”, and \emph{ŝip}, “a boat”, with the substantival termination \emph{o}.

4. If there be one negative in a clause, a second is not admissible.

5. In phrases answering the question “where?” (meaning direction), the words take the termination of the objective case; e.~g., \emph{kie\,n vi ir\,as?} “where are you going?” \emph{dom\,o\,n}, “home”; \emph{London\,o\,n}, “to London”; etc.

6. Every preposition in the international language has a definite fixed meaning. If it be necessary to employ some preposition, and it is not quite evident from the sense which it should be, the word \emph{je} is used, which has no definite meaning; for example, \emph{ĝoj\,i je tio}, “to rejoice \emph{over} it”; \emph{rid\,i je tio} “to laugh \emph{at} it”; \emph{enu\,o je la patr\,uj\,o}, “a longing \emph{for} one’s fatherland”. In every language different prepositions, sanctioned by usage, are employed in these dubious cases, in the international language, one word, \emph{je}, suffices for all. Instead of \emph{je}, the objective without a preposition may be used, when no confusion is to be feared.

7. The so-called “foreign” words, i.~e., words which the greater number of languages have derived from the same source, undergo no change in the international language, beyond conforming to its system of orthography.---Such is the rule with regard to primary words, derivatives are better formed (from the primary word) according to the rules of the international grammar: e.~g., \emph{teatr\,o}, “theater”, but \emph{teatr\,a}, “theatrical”, (not \emph{teatrical\,a}), etc.

8. The \emph{a} of the article, and the final \emph{o} of substantives, may be sometimes dropped euphoniae gratia, e.~g., \emph{de l’ mond\,o} for \emph{de la mond\,o}; \emph{Ŝiller’} for \emph{Ŝiller\,o}; in such cases an apostrophe should be substituted for the discarded vowel. 

\sectionline

% Vortaro
%
\input{vortaro.tex}

% Komento
%
\fancyhead[C]{--- \thepage ---}
\titleformat{\chapter}[display]{\centering\sansfont}{\chaptertitlename}{0pt}{\large}
\chapter*{KOMENTO PRI KOMPOSTI.}

\begin{center}
\rule{0.5\textwidth}{0.4pt}

\emph{\hfil Jen versio \laversio{} de ĉi tiu} \XeLaTeX{} \emph{aranĝaĵo.}
\end{center}

\small Kvankam mi provis rekrei la 19e-jarcentan \glqq senton\grqq{} de \emph{Unua Libro} tiel fidele kiel eble, mi ne perfekte povis imiti ĉiun tiparon, ornaman lineon, tiparan grandon, ktp.  Ankaŭ:

\begin{enumerate}

\item La rezigno per D-ro. Zamenhofo de siaj aŭtoraj rajtoj ne aperas en la angla eldono, kiun mi inspektis; tamen mi elektis aldoni ĝin al la cenzurpaĝo je la komenco, kaj aranĝis ĝin tiel, kiel ĝi estas en la eldonoj de la aliaj lingvoj, por eksponi la karakteron de la viro kaj de la internacia lingvo.

\item Tiu ĉi eldono ignoras la vorton \glqq{}kian\grqq{}, kiun D-ro. Zamenhofo ŝanĝis al \glqq{}kiam\grqq{} dum la unua jaro post la publikigo de la lingvo, kvankam la originala angla eldono de Geoghegan enhavis ambaŭ formojn; kaj same, por la aliaj korelativoj (tabelvortoj) kiun estis ŝanĝita el \glqq{}-an\grqq{} al \glqq{}-am\grqq{}.

\item Mi elektis komposti la libron por A5 papergrando, kiel la plej proksima al la originala.  Tio precipe influas la \emph{Vortaron}, kiu uzis faldatan paĝon en la originala, kiu estas tre pli granda ol la aliaj paĝoj de la libro.

\item Por anglaj vortoj kaj frazoj, mi uzis anglajn citilojn (``kiel ĉi tiuj''), anstataŭ la germanaj (\glqq{}kiel ĉi tiuj\grqq{}) de la originala.  Mi konservis la germanajn citilojn por esperantaj kaj germanaj vortoj kaj frazoj, kaj por la nomo de la lingvo Volap\"{u}k.

\item Mi elektis konservi la komecan signeton (\,), el la ĉefa teksto, kiu signifas paŭzon inter esperantaj morfemoj, kaj uzi ĝin por la \emph{Vortaro}, kiu uzis rektajn apostrofojn ({\relsize{-2.5}{$'$}}) en la originala. Tamen, mi forigis la signon post ĉiu kapvorto en la \emph{Vortaro}, kaj konservis ĝin nur, kiam estas uzita por ekzemploj.

\item Mi silente korektis malgravajn, evidentajn mistajpaĵojn, ekzemple la misliterumado de ``certain'' kiel ``certian'' ĉe la malsupro de paĝo 12 de la originalo.

\item Mi uzas modernajn nombritajn piednotojn, anstataŭ la nekonsenkvencaj ``(*)'' kaj ``*)'' el la originala.

\end{enumerate}

Mi dankas Gene-on KEYES, kies antaŭa PDF de \emph{Unua Libro} estas uzita, kiel la tajpata fundamento de la teksto.  Ĉiuj eraroj aŭ nesufiĉecoj, tamen, estas la miaj.

\vspace{1ex}

{\setlength{\parindent}{0em}
Shawn KNIGHT (angle elparolata \emph{ŝan najt})\\
\hodiau}

\newpage
\thispagestyle{empty}
\vspace*{\fill}
\begin{center}
Ĉi tiu verko estas permesita per Creative Commons \\
Attribution-NonCommercial-ShareAlike 4.0 \\
Internacia Permesilo.

La originala verko de D-ro Zamenhof \\
estas senkopirajta.\\[1ex]

\ccbyncsa\\[1ex]

This work is licensed under a Creative Commons \\
Attribution-NonCommercial-ShareAlike 4.0 \\
International License.

The original work by Dr. Zamenhof \\
is in the public domain.
\vspace*{\fill}
\end{center}
\end{document}
